Wealth dynamics in the presence of network structure and primitive cooperation

### 2019-07-05 manuscript/draftcssa19.tex
Add plug for computational simulation; wordsmith.

As mentioned earlier, macroeconomic models are ill-suited to the study of collective phenomena at an intermediate level. Hence, alternative explanations of aggregate phenomena that match the expressiveness of economic models are required. Analytical sociology~\cite{ch1as_hdbk}, with its emphasis on explanation of collective emergent phenomena using mathematically formulated social mechanisms~\cite{ch2as_hdbk,ch11as_hdbk} anchored at the individual level, is an ideal candidate for this purpose.

@@

The phenomenon under scrutiny is the emergence and evolution of economic inequality in societies with non-trivial social structures. The model we use to answer questions surrounding this phenomena must possess a dynamic model of wealth accumulation, a model of social structure; and a suitably useful measure of inequality. The dynamics are modeled as Brownian noise driven linear dynamical systems, here constant growth rate; the structure is modeled by networks, here random graphs; the measure of inequality used is Gini coefficients. Although both the dynamical system and the network model are quiet well understood, the precise interplay of structure and dynamics produces interesting emergent wealth distributions. To the best of our knowledge, this specific combination of network structure and social dynamics has not been discussed in the computational social science or mathematical sociology literature and consider this the primary contribution of this work.

Apart from enabling interactions between social actors, social structures like institutions also shape the form of cooperation and coordination mechanisms. The institutions can take the form of economic institutions, like banks and cooperatives; or the form of norms, like resource sharing practices in societies. The original SriMil model~\cite{srimil}consists of simple resource pooling arrangement where aggregates of agents pool their excess wealth in a common institution called protoinstitutions, agreeing to provide this saved resource to individual agents in times of need. The model of resource pooling used in this work is identical to the SriMil model.

While the analysis to be presented in later sections focus only on homogeneous agents with simple drift-diffusion dynamics on Erd\H{o}s-R\'{e}nyi network models, space constraints prevent us from repeating the analysis on other standard \textit{textbook} networks like scale free and small world networks. We also address more empirically grounded models in the concluding section of this paper.

In the next section, we discuss the mathematical formulation of the model. In section 3, we present the analysis of our simulation experiments, summarize key findings and discuss why we chose \texttt{julia} programming language~\cite{Julia-2017} for our implementation. In section 4, we discuss the limitations of our simple models, extensions to dynamics and networks more expressive than the ones presented here, and planned future work.

@@

\bibliography{ref_cssa19}
\end{document}

>>>>

As mentioned earlier, macroeconomic models are ill-suited to the study of collective phenomena at an intermediate level. Hence, alternative explanations of aggregate phenomena that match the expressiveness of economic models are required. Analytical sociology~\cite{ch1as_hdbk}, with its emphasis on explanation of collective emergent phenomena using mathematically formulated social mechanisms~\cite{ch2as_hdbk,ch11as_hdbk} anchored at the individual level, is an ideal candidate for this purpose. It is particularly powerful when combined with computational simulation, because when mathematically-formulated models reach even a modest level of complexity, they often become analytically intractable. Simulation \textit{in silico} can yield approximate results for these more complex scenarios, which supplement the exact results reached by traditional methods.

@@

% SD: In light of recent discussions, we may want to downplay Gini and up-play the robustness of agents to exogenous shocks.
The phenomenon under scrutiny is the emergence and evolution of economic inequality in societies with non-trivial social structures. The model we use to answer questions surrounding this phenomenon must possess a dynamic model of wealth accumulation, a model of social structure, and a suitably useful measure of inequality. The dynamics are modeled as Brownian noise driven linear dynamical systems, here constant growth rate; the structure is modeled by networks, here random graphs; the measure of inequality used is Gini coefficients. Although both the dynamical system and the network model are quite well understood, the precise interplay of structure and dynamics produces interesting emergent wealth distributions. To the best of our knowledge, this specific combination of network structure and social dynamics has not been discussed in the computational social science or mathematical sociology literature, and we consider this the primary contribution of this work.

Apart from enabling interactions between social actors, social structures like institutions also shape the form of cooperation and coordination mechanisms. The institutions can take the form of economic institutions, like banks and cooperatives; or the form of norms, like resource-sharing practices in societies. The original SriMil model~\cite{srimil}consists of a simple resource pooling arrangement where aggregates of agents pool their excess wealth in a common institution called ``proto-institutions,'' agreeing to provide this saved resource to individual agents in times of need. The model of resource pooling used in this work is identical to the SriMil model.

The analysis to be presented in later sections focuses only on homogeneous agents with simple drift-diffusion dynamics on Erd\H{o}s-R\'{e}nyi network models; space constraints unfortunately prevent us from repeating the analysis on other standard \textit{textbook} networks like scale-free and small-world networks. We also address more empirically grounded models in the concluding section of this paper.

In the next section, we discuss the mathematical formulation of the model. In section 3, we present the analysis of our simulation experiments, summarize key findings and discuss why we chose the \texttt{julia} programming language~\cite{Julia-2017} for our implementation. In section 4, we discuss the limitations of our simple models, extensions to dynamics and networks more expressive than the ones presented here, and planned future work.

@@

\bibliography{ref_cssa19}
\end{document}
% this line is for Stephen's purposes only:
% vim:textwidth=99999
